Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-15 main.tex
some more defs

\newcommand{\gA}{\graph{A}}
\newcommand{\gB}{\graph{B}}
\newcommand{\gC}{\graph{C}}
\newcommand{\gG}{\graph{G}}
\newcommand{\gH}{\graph{H}}
\newcommand{\gN}{\graph{N}}

@@

\newcommand{\cons}[1]{\{#1\}}
\newcommand{\Scons}[1]{S\cons{#1}}
\newcommand{\conhole}{\cons{~}}
\newcommand{\Sconhole}{S\conhole}

@@

If a derivation $\Deri$ contains instances of the rules $\mfaD$ and
$\mexD$ we can no longer naively define the rectification
$\rectif\Deri$ as in the previous section for $\FOMLS$, as these two
rules cannot be applied if premise and conclusion are rectified. For
this reason we define here rectified versions $\rectif\mfaD$ and
$\rectif\mexD$.

\todo{define rules $\rectif\mfaD,\rectif\mexD$}

\todo{define $\mapof\Deri$ (underlying map doing the flowgraph thing)}

\begin{lemma}
  Let $\vlderivation{
    \vlde{\set{\wlD,\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\Deri}{B}{
      \vlhy{A}}}$ be a rectified derivation. Then the map
  $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ is a skew bifibration.
\end{lemma}

>>>>

\newcommand{\vgraphof}[1]{V_{\graphof{#1}}}
\newcommand{\gA}{\graph{A}}
\newcommand{\gB}{\graph{B}}
\newcommand{\gC}{\graph{C}}
\newcommand{\gG}{\graph{G}}
\newcommand{\gH}{\graph{H}}
\newcommand{\gN}{\graph{N}}

@@

\newcommand{\cons}[1]{\{#1\}}
\newcommand{\Scons}[1]{S\cons{#1}}
\newcommand{\conhole}{\cons{\cdot}}
\newcommand{\Sconhole}{S\conhole}

@@

If a derivation $\Deri$ contains instances of the rules $\mfaD$ and
$\mexD$ we can no longer naively define the rectification
$\rectif\Deri$ as in the previous section for $\FOMLS$, as these two
rules cannot be applied if premise and conclusion are rectified. For
this reason we define here rectified versions $\rectif\mfaD$ and
$\rectif\mexD$, show below:
\begin{equation*}
  \vlinf{\rectif\mfaD}{}{\Scons{\forall x.(Ax\vlor Bx)}}{\Scons{(\forall y.Ay)\vlor(\forall z.Bz)}}
  \qquad
  \vlinf{\rectif\mexD}{}{\Scons{\exists x.(Ax\vlor Bx)}}{\Scons{(\exists y.Ay)\vlor(\exists z.Bz)}}
\end{equation*}
Here, we use the notation $A\cdot$ for a formula $A$ with occurrences
of a placeholder $\cdot$ for a variable. Then $Ax$ stands for the
results of replacing that placeholder with $x$, and also indicating
that $x$ must not occur in $A\cdot$. Then $\forall x.Ax$ and $\forall
y.Ay$ are the same formula modulo renaming of the bound variable bound
by the outermost $\forall$-quantifier. In the rules $\rectif\mfaD$ and
$\rectif\mexD$ we also demand that the variables $x$, $y$, and $z$ do
not occur in the context $\Sconhole$.

For each instance $\vlinf{\rr}{}{P}{Q}$ of an inference rule in
$\set{\wlD,\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can
define the \bfit{induced map} $\mapof{\rr}\colon\vgraphof
Q\to\vgraphof P$ which acts as the identity for
$\rr\in\set{\me,\fequ}$ and as the canonical injection for
$\rr\in\set{\wlD,\wrD}$. For $\rr=\acD$ it maps the vertices
corresponding to the two atoms in the premise to vertex of the
contracted atom in the conclusion, and for
$\rr\in\set{\rectif\mfaD,\rectif\mexD}$ it maps the two vertices
corresponding to the quantifiers in the premise to the one in the
conclusion (as acts as the identity on all other vertices). For a
derivation $\Deri$ in
$\set{\wlD,\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we define
the \bfit{induced map} $\mapof\Deri$ as the composition of the induced
maps of the rule instances in $\Deri$.

\todo{example}
\begin{lemma}
  Let $\vlderivation{
    \vlde{\set{\wlD,\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}}{\Deri}{B}{
      \vlhy{A}}}$ be a rectified derivation. Then the induced map
  $\mapof\Deri\colon\graphof{A}\to\graphof{B}$ is a skew bifibration.
\end{lemma}
